---
title: "week7_assignment7"
author: "Jayanth0572"
date: "October 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> # Reading files in multiple formats(XML, HTML and JSON) into R
  - XML and HTML files require XML library and for JSON file, jsonlite library was used. There were also other options, but these were pretty straight forward.
  - When we create dataframes, the datafrmaes for XML and HTML imports looked similar. The column types were all factors. For JSON import, the column with array input was read as a list instead of reading as a string with the values seperated by some delimiter. rest of the columns types were character.


# Reading XML data into R
- Requires XML library import
- "xmlToDataFrame"" function was used to create dataframe by directly reading from the input XML file
- The column types are all factors

```{r}

library(XML)

# creating dataframe from the input xml using xmlToDataFrame function
books_xml_df <- xmlToDataFrame("books.xml")
#printing the dataframe
books_xml_df

# class is dataframe and type is list
(class(books_xml_df))
(typeof(books_xml_df))

```



# Reading HTML data into R
- Requires XML library import
- "readHTMLTable"" function was used to read the input HTML file as a table into R. Te result will be a ist
- Then the list was converted to a dataframe
- The column types are all factors

```{r}

library(XML)

# Reading HTML file into R using xmlParse function
html_obj <-readHTMLTable("books.html", header = T)
# Assigning name to the list
names(html_obj) <- "books"
#printing the list
html_obj

# Converting list to a dataframe
books_html_df <- html_obj$'books'
#printing the dataframe
books_html_df

# class is dataframe and type is list
(class(books_html_df))
(typeof(books_html_df))

```


# Reading JSON data into R
- Requires jsonlite library import
- "fromJSON"" function was used to read the input JSON file as a table into R. Te result will be a ist
- Then the list was converted to a dataframe
- The column types are combination of character and list
> Array we used for authors was imported as a list

```{r}

library(jsonlite)

# Getting json data into R. fromJSON function imports the jasaon data as list
json_obj <- fromJSON("books.json")
#printing the list
json_obj

# converitng list to a dataframe.
# Note1: Notice, the Authors column is a list
books_json_df <- json_obj$'books'
#printing the dataframe
books_json_df

# Authors by row sequence
(books_json_df$Authors)

# class is dataframe and type is list
class(books_json_df)
typeof(books_json_df)

```


